Throughput Prediction on 60 GHz Devices for Latency Sensitive Applications

### 2021-01-21 paper.tex
changes

In this paper, we show that there is a clear region in the field-of-view (FoV) outside which 60 GHz does not work. We also show that within this FoV, throughput may not always be stable as has been previously claimed. Using neural networks, we predict the throughput for the 60 GHz link at different timescales, from 10 ms (suitable for VR) up to 2 s (suitable for ABR streaming). We show that we can very accurately predict the throughput for various realistic scenarios across different time scales.

@@

Over the past few years consuming video content has become tremendously popular. As of 2019, YouTube had 2 billion logged-in monthly users~\cite{statista} and more than 73\% of adults in the US watched videos on YouTube~\cite{pew-research}. More recently, there has been an influx of novel types of multimedia content such as 360\degree~videos, volumetric videos, and VR gaming. Users also have novel ways to consume such content, such as through a head mounted display (HMD). However, such content also has a higher demand in terms of the user's quality of experience (QoE). However, video streaming is known to be challenging; the network conditions between the server and the user inevitably have variations, which could cause a user watching a 360\degree~video or playing a VR game to experience a long start up time or periods of rebuffering/stalls. Such experiences can have different consequences for different types of content. For example, a user watching a video that keeps getting stalled may just stop watching the video while a user playing a VR game might suffer from motion sickness.

To deal with such issues, adaptive bitrate (ABR) streaming has become one of the most popular techniques to stream multimedia content. Consequently, a number of adaptive streaming standards has been introduced over the years~\cite{stockhammer:mmsys2011, pantos-hls, adobe-hds, microsoft-ss}. ABR streaming monitors the network conditions and adapts the content quality to minimize the occurrences of such issues, which could drastically impact the user QoE. One of the main components of adaptive bitrate streaming is the estimation of network conditions. Traditionally, most ABR systems calculate throughput estimates and choose a quality level based on the download speeds of the past chunks and/or playback buffer occupancy~\cite{sun:sigcomm2016, spiteri:infocom2016}. More recently, the use of machine learning (ML) to select the most appropriate quality level has gained significant popularity~\cite{mao:sigcomm2017, yan:nsdi2020} and ML-based ABR algorithms have been shown to outperform traditional ones.

>>>>

In this paper, we show that there is a clear region in the
field-of-view (FoV) outside which 60 GHz does not work. We also show
that within this FoV, throughput may not always be stable as has been
previously claimed. Using neural networks, we predict the throughput
for the 60 GHz link at different timescales, from 10 ms (suitable for
VR) up to 2 s (suitable for ABR streaming). We show that we can 
accurately predict the throughput for various realistic scenarios
across different time scales.

@@

Over the past few years consuming video content has become
tremendously popular. As of 2019, YouTube had 2 billion logged-in
monthly users~\cite{statista} and more than 73\% of adults in the US
watched videos on YouTube~\cite{pew-research}. More recently, there
have been an influx of novel types of multimedia content such as
360\degree~videos, volumetric videos, and VR gaming. Users also have
novel ways to consume such content, such as through a head mounted
display (HMD). 
% However, such content also has a higher demand in terms
% of the user's quality of experience (QoE). 
However, video streaming is
known to be challenging; the network conditions between the server and
the user inevitably have variations, which could cause a user watching
a 360\degree~video or playing a VR game to experience a long start up
time or periods of rebuffering/stalls. Such experiences can have
different consequences for different types of content. For example, a
user watching a video that keeps getting stalled may just stop
watching the video while a user playing a VR game might suffer from
motion sickness.

To deal with such issues, adaptive bitrate (ABR) streaming has become
one of the most popular techniques to stream multimedia
content. Consequently, a number of adaptive streaming standards have
been introduced over the years~\cite{stockhammer:mmsys2011,
  pantos-hls, adobe-hds, microsoft-ss}. ABR streaming monitors the
network conditions and adapts the content quality to minimize the
occurrences of such issues, which could drastically impact the user
QoE. One of the main components of adaptive bitrate streaming is the
estimation of network conditions. Traditionally, most ABR systems
calculate throughput estimates and choose a quality level based on the
download speeds of the past chunks and/or playback buffer
occupancy~\cite{sun:sigcomm2016, spiteri:infocom2016}. More recently,
the use of machine learning (ML) to select the most appropriate
quality level has gained significant popularity~\cite{mao:sigcomm2017,
  yan:nsdi2020} and ML-based ABR algorithms have been shown to
outperform traditional ones.
